---
title: "EDA using group_by"
author: "Marisa Sobel"
date: "10/4/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
  )

library(tidyverse)
library(ggridges)
library(patchwork)
library(ggthemes)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

```{r data_import, cache = TRUE, collapse=TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2016-01-01",
                      date_max = "2016-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, date, month, everything())
weather_df
```

## Group by name and month 

```{r, collapse=TRUE}
weather_df %>% 
  group_by(name, month)
```

## Grouped summaries 

NOTE: group and sum will repeate the action across all groups, different number outcomes based on what is in the group

NOTE: group and mutate will repeate outcome across all groups (?)

```{r}
# group by name and month --> show number of distinct in each group
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n())

# group by month --> show number of distinct in each group
weather_df %>% 
  group_by(month) %>% 
  summarize(number = n())

# count how many times each month comes up
weather_df %>% 
  count(month)

# group by month --> number of distinct dates 
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n(),
            n_days = n_distinct(date))

# group by name and month --> number of distinct dates and the average tmax
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(), 
            mean_tmax = mean(tmax))

# same as above, but REMOVE MISSING VARS and STACKING up different summaries in the same table
weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(), 
            mean_tmax = mean(tmax, na.rm = TRUE),
            median_prcp = median(prcp, na.rm = TRUE), 
            sd_tmin = sd(tmin, na.rm = TRUE))
```

There is still a tibble! 

```{r}
month_df = 
  weather_df %>% 
  group_by(name, month) %>% 
  summarize(number = n(), 
            mean_tmax = mean(tmax, na.rm = TRUE),
            median_prcp = median(prcp, na.rm = TRUE), 
            sd_tmin = sd(tmin, na.rm = TRUE)) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point()
```

Plot two tibbles at once!

```{r, eval=FALSE}
# CODE NOT WORKING

# use geom_data argument
ggplot(data = weather_df, aes(x = month, y = mean_tmax, color = name)) +
  geom_point(alpha = .5) +
   geom_line(data = month_df, aes(x = month, y = mean_tmax))
```

Today not always best for humans...

```{r, eval=FALSE}
# CODE NOT WORKING
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) %>% 
  spread(key = month, value = mean_tmax) %>% 
  knitr::kable(digits = 3)
```

## Grouped mutate

```{r}
# 
weather_df %>% 
  group_by(name, month) %>% 
  mutate(mean_tmax = mean(tmax, na.rm = TRUE))

# mean tmax throughout whole dataset as intermediate variable to get centered tmax
# everything is not centered around zero
# how extremes relate to averages 
weather_df %>%
  group_by(name) %>%
  mutate(mean_tmax = mean(tmax, na.rm = TRUE), 
    centered_tmax = tmax - mean(tmax)) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() 
```

## Window Functions 

Use ranks

```{r}
# use min_rank over rank --> will give smallest ranking value 
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(tmax))

# filter < 2 --> finds the coldest day
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(tmax)) %>% 
  filter(tmax_rank < 2)

# descending order
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(desc(tmax)))
```

Use lags 

```{r}
weather_df %>% 
  group_by(name) %>%
  mutate(lag_tmax = tmax - lag(tmax))
```

## Look at this...

```{r}
# simple function for fairly complicated questions
weather_df %>% 
  group_by(name) %>%
  mutate(one_day_change = tmax - lag(tmax)) %>% 
  summarise(sd_one_day_change = sd(one_day_change, na.rm = TRUE)) %>% 
  kable()
```




