---
title: "Visualization Pt 1"
author: "Marisa Sobel"
date: "9/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)

```

```{r data_import, cache = TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Start a plot

Look at `t max` vs `t min`

Blank plot...

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

Scatterplot 

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

Scatterplot `CentralPark_NY` only, using pipes

```{r, collapse=TRUE}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  ggplot(aes( x = tmin, y = tmax)) +
  geom_point()
```

Name and save

```{r, collapse=TRUE}
weather_plot = 
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

## More plot options

Add an aesthetic - colors different by place 

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```

Add a geom

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .4) + 
  # alpha blending level for transparency 
  geom_smooth(se = FALSE)
```

putting `color = name` in the main `ggplot` object applies it to all geoms (globally rather than locally)

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .4) + 
  geom_smooth(se = FALSE)
```

Facetting...

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(alpha = .4) + 
  geom_smooth(se = FALSE) +
  facet_grid( ~ name) 
  # separate panels on `name` variable 
```

A more intersting plot

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) + 
  geom_smooth(se = FALSE) +
  facet_grid( ~ name)
```

`geom_point` --> data points
`geom_line` --> connects the dots





