---
title: "Visualization 2"
author: "Marisa Sobel"
date: "10/2/2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
  )

library(tidyverse)
library(ggridges)
library(patchwork)
library(ggthemes)
```

```{r data_import, cache = TRUE}
library(rnoaa)

weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```


## Start a plot

First scatterplot

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
    )
```

Tick marks and labels ...

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
    ) + 
  scale_x_continuous(breaks = c(-15, 0, 15), 
                     labels = c("-15ºC", "0", "15"),   # option + 0 = º
                     limits = c(-20, 30)) + 
  scale_y_continuous(trans = "sqrt", 
                     position = "right")
```

## Colors and themes

Adjust color

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  scale_color_hue(name = "Location", 
                  h = c(100, 300),         # range of hues 
                  l = 50)    
```

Viridis package 

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE   # makes discrete from continuous scale 
  ) + 
  theme(legend.position = "bottom")
```

## Themes

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE   # makes discrete from continuous scale 
  ) + 
  theme_bw() +        # ORDER MATTERS! - set theme first, then modify!
  theme(legend.position = "bottom")
```

## Practice!

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = prcp), alpha = .75) + 
  labs(
    title = "Temperature Plot", 
    x = "Date", 
    y = "Max Daily Temp (ºC)",
    caption = "Data from the rnoaa package"
  ) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```


## Arguments to `geom_*`

```{r, collapse=TRUE}
centralpark_df = weather_df %>%  filter(name == "CentralPark_NY")
waikiki_df = weather_df %>%  filter(name == "Waikiki_HA")

ggplot(waikiki_df, aes(x = date, y = tmax, color = name)) +
  geom_point() + 
  geom_point(data = centralpark_df)

```

## patchwork 

Separate the plots...

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) + 
  facet_grid( ~ name) +
  geom_point(aes(size = prcp), alpha = .75) + 
  labs(
    title = "Temperature Plot", 
    x = "Date", 
    y = "Max Daily Temp (ºC)",
    caption = "Data from the rnoaa package"
  ) +
  viridis::scale_color_viridis(discrete = TRUE) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

...but sometime might want to make distinctly different plots!

```{r, collapse=TRUE}
tmax_tmin_p = ggplot(weather_df, aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

# stitching together fundamentally different plots
(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

## Data manipulation 

factors...they show up in the order of that factor

```{r, collapse=TRUE}
ggplot(weather_df, aes(x = date, y = tmax, fill = name)) + 
  geom_violin()
```

Ordering factors that explicitly choose, rather than the way it is

```{r, collapse=TRUE}
# select the way you want the levels, "by hand"
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")

# reorder name by tmax 
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

## Advanced tidying...

* tmax and tmin are different observations 
* focus only on vars needed
* want to go wide --> long

```{r, collapse=TRUE}
weather_df %>%
  select(name, tmax, tmin) %>% 
  gather(key = observation, value = temp, tmax:tmin) %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```




